---
  title: "Query CDISC ADAE"
  output: github_document
---

  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(here)
library(RSQLite)
library(dbplyr)  ## may need to update dbplyr 
library(DBI)
```


Load example CDISC data from [https://github.com/phuse-org/phuse-scripts](https://github.com/phuse-org/phuse-scripts).

```{r}
adsl <- haven::read_xpt(here("data","adsl.xpt"))
adae <- haven::read_xpt(here("data","adae.xpt"))
```

Calculate the number of patients by treatment in the safety population (SAFFL). NOTE: All of these steps to generate summary statistics could be performed within a database using SQL. 

```{r}
bigN <- adae %>%
  filter(SAFFL == "Y") %>% ## SAF population
  group_by(TRTA) %>%
  summarise(N = n()) %>%
  mutate(N = N)

```

Calculate the number of patients by treatment assigned by adverse event (dictionary dervived term). 

```{r}
smalln <- adae %>% 
  filter(SAFFL == "Y") %>% ## SAF population
  group_by(AEDECOD, TRTA) %>% # group by PT - disctionary dervived term and trt
  summarise(n = n()) %>%
  mutate(n = n) 

```

Finally, calculate the percentage of events by derived term i.e. the treatment emergent adverse events. 

```{r}
teae <- bigN %>%
  left_join(smalln) %>%
  mutate(inc = n / N, 
         per = n / N * 100) %>%
  arrange(TRTA, AEDECOD) %>% 
  glimpse()
```


## Database


```{r}
ARD <- teae %>%
  mutate(
    ANALID = row_number(),
    STUDYID = "1",
    UANALID = paste0(STUDYID, ".", ANALID),
    ANALFLG = "SAFFL",
    TRTVAR = "TRTA",
    TRTVAL = TRTA,
    SUBVAR = "",
    SUBVAL = "",
    AVISTN = "EOS",
    ANALMETH = "Summary statistics",
    PARAM	= "treatment emergent adverse event (dictionary dervived term)",
    PARAMCD	= "TEAE" ,
    ROWCAT1 = AEDECOD,
    ANALTYP1 = "DESCRIPTIVE",
    ANALTYP2 = "",
    ANALDESCRIPTION = "Descriptive summary of treatment emergent adverse events by treatment."
  ) %>%
  select(-AEDECOD, -TRTA) %>%
  gather(
    "STAT",
    "STATVAL",
    -UANALID,
    -STUDYID,
    -ANALID,
    -ANALFLG,
    -TRTVAR,
    -TRTVAL,
    -SUBVAR,
    -SUBVAL,
    -AVISTN,
    -PARAM,
    -PARAMCD,
    -ROWCAT1,
    -ANALTYP1,
    -ANALTYP2,
    -ANALMETH,
    -ANALDESCRIPTION
  ) %>%
  arrange(STUDYID, ANALID, STAT) %>%
  glimpse()
```


Write to a database the results. 




```{r}
mydb <- dbConnect(RSQLite::SQLite(), "")
dbWriteTable(mydb, "ARD", ARD)
```

```{r}
dbListTables(mydb)
```

```{r}
dbListFields(mydb, "ARD") 
```

Query the database for AEs with an incidence over 5%. 


```{r}
dbdata <- tbl(mydb, "ARD")
dbdata %>%
  filter(ANALTYP1 == "DESCRIPTIVE" &  PARAMCD == "TEAE" & STAT == "inc" && STATVAL > 0.05 ) %>%
  select(ANALID, TRTVAL, ROWCAT1, STAT, STATVAL) 
```

Disconnect the database. 

```{r}
dbDisconnect(mydb)
```

## Session info
```{r}
sessionInfo()
```
