---
title: "Example of reading & writing to ARD"
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is an example of: 

* selecting (i.e. a query) from an analysis results database an endpoint from three clinical studies for the PBO arm,
* generating a historical control,
* writing the estimates from the MAP prior back to the ARD database. 

## Set up
Load the required packages and path to a CSV file containing the analysis dataset using a mock data model. 

TODO: The model is not yet broken down by entity-relationship. 

```{r setup}
library(knitr)
library(tidyverse)
library(RBesT)
library(RSQLite)  ## require an updated version fix column order bug when appending 
library(DBI)
path_to_data <- "./data/ARD-example.csv"
ARD <- read.csv(path_to_data)
```

## Load dataset in to SQL database

Use RSQLite as a proof of concent. 

Note: need to update the version of RSQlite to address an issue with column ordering when appending back to the database. 

First step, set up sqlite database. 

```{r}
mydb <- dbConnect(RSQLite::SQLite(), "my-db.sqlite")
dbDisconnect(mydb)
unlink("my-db.sqlite")
mydb <- dbConnect(RSQLite::SQLite(), "")
dbDisconnect(mydb)
```

Insert the analysis result data.frame as a table in the database. This table has a lot of duplication that can be simplified by fully modelling the entity-relationships. 

```{r}
mydb <- dbConnect(RSQLite::SQLite(), "")
dbWriteTable(mydb, "ARD", ARD)
```

Display the column names in the table as a sanity check. 

```{r}
dbListTables(mydb)
```


Also display the first five rows of the new ARD table. 

```{r}
dbGetQuery(mydb, 'SELECT * FROM ARD LIMIT 5')
```

## Query the database

Now we can query database. 

We want to query for studies that have: 

* an outcome "PCE" which is short for "Primary composite endpoint", 
* the analysis is "descriptive" (ANALYTYP1 = DESCRIPTIVE""),
* and a placebo treatment arm (TRTVAL = PBO),
* return the selected data and write to dataframe,
* and transform ready for RBesT,
* display a glimpse of the df structure to ensure this worked. 

```{r}
df <- dbGetQuery(mydb, 'SELECT STUDYID, TRTVAL, STAT, STATVAL FROM ARD WHERE PARAMCD = "PCE" AND ANALTYP1 = "DESCRIPTIVE" AND TRTVAL = "PBO" ') %>%
  group_by(STUDYID, TRTVAL, STAT) %>%
   mutate(id = row_number()) %>% 
   gather(groupname, value, -id, -STUDYID, -TRTVAL, -STAT) %>%
   spread(STAT, value) %>%
   select(TRTVAL, N, n) %>%  
   glimpse()
  
```

## MAP prior of three studies using RBesT

We run the placebo arm through RBest to generate a MAP prior. Note: we are just using the default settings without any thought to make sure the data is in the correct format. 

```{r}
set.seed(34563)
map_mcmc <- gMAP(cbind(n, N-n) ~ 1 | STUDYID,
                 data=df,
                 tau.dist="HalfNormal",
                 tau.prior=1,
                 beta.prior=2,
                 family=binomial)
```


Print the output of the model. 

```{r}
print(map_mcmc)
```

Check that a graphical representation of model checks is available
```{r}
pl <- plot(map_mcmc)
```

Check that a number of plots are immediately defined
```{r}
names(pl)
```

Display the forest plot with model estimates 
```{r}
print(pl$forest_model)
```

## Arrange data and write back to the db 

Now we need to arrange the data into a format that can be inserted back in to the database. 

To do so we:

* Map the output to datatable ARD schema (NOTE: this is very manual at the moment),
* gather the statistics from the three mcmc chains into seperate analyes
* give each analysis its own unique id
* indicate that the analysis comes from three rather than one study. 



```{r}
map <- as_tibble(map_mcmc$theta.strat) %>%
  mutate(
    ANALID = row_number(),
    STUDYID = "123",
    UANALID = paste0(STUDYID, ".", ANALID),
    ANALFLG = "FAS",
    TRTVAR = "treatment",
    TRTVAL = "PBO",
    SUBVAR = "",
    SUBVAL = "",
    AVISTN = "EOS",
    ANALMETH = "MAP prior",
    PARAM	= "Primary composite outcome: CVD or FHFH",
    PARAMCD	= "PCE",
    ROWCAT1 = "ROWCAT1",
    ANALTYP1 = "PREDICTIVE",
    ANALTYP2 = "",
    ANALDESCRIPTION = "RBesT MAP Prior PBO from study 1, 2 and 3"
  ) %>%
  gather(
    "STAT",
    "STATVAL",-UANALID,
    -STUDYID,
    -ANALID,
    -ANALFLG,
    -TRTVAR,
    -TRTVAL,
    -SUBVAR,
    -SUBVAL,
    -AVISTN,
    -PARAM,
    -PARAMCD,
    -ROWCAT1,
    -ANALTYP1,
    -ANALTYP2,
    -ANALMETH,
    -ANALDESCRIPTION
  ) %>%
  glimpse()

```

Check that the data.frame maps to the ARD schema by displaying the columns in table ARD. 

```{r}
dbListFields(mydb, "ARD") 
```

## Append to database

Now we write back to the database appending to the table. 


```{r}
dbWriteTable(mydb, "ARD", map, append = TRUE)

```


We then check that the data from the new analysis is there. 

```{r}
dbGetQuery(mydb, 'SELECT * FROM ARD WHERE PARAMCD = "PCE" AND ANALTYP1 = "PREDICTIVE" AND TRTVAL = "PBO" ')
```

Disconnect the database. 

```{r}
dbDisconnect(mydb)
```


## Session info
```{r}
sessionInfo()
```


